---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DESeq2)
```

```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"

# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)
```

```{r}
countData = read.csv(countFile, row.names=1)
head(countData)
```

```{r}
# Note we need to remove the odd first $length col
countData <- as.matrix(countData[,-1])
head(countData)
```

The length column in countdata is going to cause probs with the requires matching to the metadata 



```{r}
colnames(countData)
```

```{r}
rownames(colData)
```




```{r}
all(colnames(countData)==rownames(colData))
```


```{r}
#test how the all() 
all(c(T,F,T))
```

We want to remove the zero genes that have 0 count values in all expreiments.
```{r}
countData=countData[rowSums(countData)!=0,]
head(countData)
```


#Deseq2
```{r}
library(DESeq2)
```


```{r}
#setup the obj w/ out data in the way DESeq wants it
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
#Run the analysis
dds = DESeq(dds)
```

Get our results 
```{r}
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
```


```{r}
res
```


make a volcano plot
```{r}
plot( res$log2FoldChange, -log(res$padj) )
```


```{r}
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
mycols[abs(res$log2FoldChange)>2]<-"blue"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```

##Add gene symbol and 
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
#we can translate btw all the following database ID system
columns(org.Hs.eg.db)
```


```{r}
res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(countData), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")
res
```

```{r}
res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")
```

#Pathway analysis
Here we are going to use the gage package for pathway analysis. Once we have a list of enriched pathways, we're going to use the pathview package to draw pathway diagrams, shading the molecules in the pathway by their degree of up/down-regulation.
```{r}
library(pathview)
library(gage)
library(gageData)
```

```{r}
data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```


```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```
greater =up reg
less =down red

```{r}
# Look at the first few down (less) pathways
head(keggres$less)
```


```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```



```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```



```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```
```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```



#gene ontoloogy (GO)
We can also do a similar procedure with gene ontology. Similar to above, go.sets.hs has all GO terms. go.subs.hs is a named list containing indexes for the BP, CC, and MF ontologies. Let’s focus on BP (a.k.a Biological Process) here.

```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```






