---
title: "class 11: structural bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##The PDB database for biomoleculer structure data

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
you determine what proportion of structures are protein?

download CSV (accessible from “Analyze” -> “PDB Statistics” >“by Experimental Method and Molecular Type”)
```{r}
#read csv
data<- read.csv("Data Export Summary.csv")
data
```


total number of enteries
```{r}
total<-sum(data$Total)

```

proportion from each method
```{r}
(data$Total/sum(data$Total))*100
```

propotion that are protein
```{r}
round(sum(data$Proteins)/sum(data$Total)*100,2)
```

Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

Direction
1.Go to PDB and Download the PDB file
2. Open VMD and go to the main window medium(there is three: big, medium, and small)
3. Use file broswer to open desired file
4. To change graphics go to VMD and click graphics
5. Change drawing methods

other things you can change
coloring method 
selecteed atoms is the atoms you ant to change
ex: water
not protein and not water
create rep: makes a new file for you to change w/o affecting the other



Q3: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule
in this structure?
Because the water molecules are too small for the resolution
Q4: There is a conserved water molecule in the binding site. Can you identify this water
molecule? What residue number does this water molecule have (see note below)?
there are four resid192


```{r}
library(bio3d)

pdb<-read.pdb("1hsg")
```








```{r}
ligand<-
```






