---
title: "Class 7 R functions and packages"
author: "Tiffany"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,10,5,NA,6))
```


```{r},eval= FALSE
rescale2(c(1:10, "barry"))
```

```{r}, eval= FALSE
rescale(c(1:10, "barry"))
```

##A new function called both_na()
Write a function to find where there are NA elements in two input vectors.

First make some simple input where I know the answer
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

Looked online and found the **is.na()** function
```{r}
is.na(x)
```

```{r}
which(is.na(x))
```
and the **which()** function tells me where the TRUE values are
```{r}
which(is.na(y))
```


```{r}
is.na(x)
is.na(y)
```
The AND function requires two input TRUE to give a TRUE output
```{r}
is.na(x)& is.na(y)
```

Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have.This is my working snippet!!
```{r}
sum(is.na(x)& is.na(y))
```

```{r}
sum(c(TRUE,TRUE,TRUE,FALSE))
```


Now turn it into a function

```{r}
both_na<-function(x,y) {
sum(is.na(x)& is.na(y))
}
```

```{r}
both_na(x,y)
```


```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

```

```{r}
both_na(x, y1)
```

```{r}
# What will this return?
both_na(x, y2)
```

```{r}
x2 <-c(NA,NA)
```

```{r}
both_na(x2,y2)
```


```{r}
plot(1:10, col= c("red","blue","green"))
```


```{r}
x<- c(1, NA, NA)
y3<- c(1, NA, NA, NA, NA, NA, NA)
both_na(x,y3)
```


```{r}
x3<- c(1, NA, NA, 1, NA, NA, 1)
y3<- c(1, NA, NA, NA, NA, NA, NA)
both_na(x,y3)
```

```{r}
length(x)
length(y3)
```


```{r}
both_na2<-function(x,y) {
if (length(x)!=length(y)){
  stop("Inputs x and y should be the same length!")
}
sum(is.na(x)& is.na(y))
}
```


```{r}, eval= FALSE
both_na2(x,y3)
```

```{r}
s1<-c(100,100,100,100,100,100,100,90)
s2<-c(100, NA, 90,90,90,90,97,80)
s1[-which.min(s1)]
s2[-which.min(s2)]
```


```{r}
mean(s1)
mean(s2)
```

This is our working snippet
```{r}
s2[-which.min(s2)]
mean(s2[-which.min(s2)], na.rm=TRUE)
```

This is our working code
```{r}
# there was another equation that i didn't get
```





```{r}
grade<- function(x){
  if (any(is.na(x))) {
    warning("Student is missing HW")
  }
sum(x[-which.min(x)], na.rm=TRUE)/ (length=(x)-1)
}
```

```{r}
grade(s2)
```




```{r}
url<-("https://tinyurl.com/gradeinput")
hw<- read.csv(url)
```



```{r},eval= FALSE
install.packages("datapasta")
```

tibble::tribble(
  ~No.,                     ~Location, ~Station.ID,    ~Amount,
     1,       "Nador-Aroui (Morocco)",       60340,   "318 mm",
     2,     "Dakhla (Western Sahara)",       60096, "205.6 mm",
     3,    "Matlapa, S.l.p. (Mexico)",       76585,   "194 mm",
     4,            "Alapuzha (India)",       43352,   "182 mm",
     5, "Cochin / Willingdon (India)",       43353,   "166 mm",
     6,     "Tampico, Tamps (Mexico)",       76548, "160.1 mm",
     7,            "Karaikal (India)",       43346, "124.4 mm"
  )























